% CHAPTER 4 — DEVELOPMENT AND PROTOTYPE SOLUTION

\chapter{Development and Prototype Solution}

% ─────────────────────────────────────────────────────────────────────────────
\section{Introduction}
% ─────────────────────────────────────────────────────────────────────────────

This chapter presents the development phase of the BodyQ mobile application,
from the technical decisions taken during the design phase through to the
complete working prototype. The objective of this phase was to transform the
requirements defined in the previous chapters into a fully functional,
AI-powered fitness application that runs on both iOS and Android devices.

The development of BodyQ was guided by three core principles:

\begin{itemize}
    \item \textbf{Personalization:} Every feature — from the workout plan to
    the daily meal suggestions — must adapt to each individual user's profile,
    goals, and physical condition.

    \item \textbf{Intelligence:} The application must go beyond static content
    by integrating real artificial intelligence that understands the user,
    learns from their activity, and responds meaningfully.

    \item \textbf{Usability:} A fitness app is only effective if users engage
    with it daily. Every screen, interaction, and flow was designed to be
    intuitive, fast, and visually appealing.
\end{itemize}

The prototype was developed as a cross-platform mobile application using
React Native with Expo, connected to a Supabase backend, and powered by two
AI providers: Groq (for conversational AI and plan generation) and Google
Gemini (as an alternative plan generation engine). The result is a complete,
deployable prototype covering user authentication, onboarding, AI workout and
meal planning, real-time tracking, analytics, community features, and an AI
coaching assistant named Yara.

% ─────────────────────────────────────────────────────────────────────────────
\section{Development Details}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Technology Stack}

The choice of technology was driven by the need to deliver a high-quality
cross-platform application quickly, with a scalable backend and built-in AI
capabilities. Table~\ref{tab:techstack} summarizes the full technology stack.

\begin{table}[h!]
\centering
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{|p{3.5cm}|p{4cm}|p{6cm}|}
\hline
\textbf{Component} & \textbf{Technology} & \textbf{Justification} \\
\hline
Mobile Framework & React Native + Expo &
    Single codebase for iOS and Android; Expo accelerates development with
    managed workflow \\
\hline
Language & JavaScript + TypeScript &
    TypeScript used for all service and model layers to enforce type safety \\
\hline
Backend \& Database & Supabase (PostgreSQL) &
    Provides Auth, Database, Storage, and serverless Edge Functions in one
    platform \\
\hline
AI Chat \& Planning & Groq API (LLaMA models) &
    Ultra-fast inference; \texttt{llama-3.1-8b-instant} for chat,
    \texttt{llama-3.3-70b-versatile} for plan generation \\
\hline
AI Plan Generation & Google Gemini 2.0 Flash &
    Used in the onboarding plan flow; generous free tier (1M tokens/min) \\
\hline
Local State & React Context API &
    Lightweight global state without Redux overhead \\
\hline
Local Persistence & AsyncStorage &
    Offline schedule storage and completion tracking \\
\hline
Navigation & React Navigation &
    Industry standard for React Native screen navigation \\
\hline
\end{tabular}
\caption{Application Technology Stack}
\label{tab:techstack}
\end{table}

\subsection{Application Architecture}

The application follows a layered architecture that cleanly separates
responsibilities across five layers, as illustrated in
Figure~\ref{fig:architecture}.

\begin{itemize}
    \item \textbf{Screens} — Full-page views that compose components and call
    hooks for their data logic.
    \item \textbf{Components} — Reusable UI pieces such as cards, progress
    bars, overlays, and modals.
    \item \textbf{Hooks} — Custom React hooks encapsulating business logic,
    one per feature domain (\texttt{useWorkout}, \texttt{useNutrition},
    \texttt{useInsights}, etc.).
    \item \textbf{Services} — All API calls and database operations
    (\texttt{workoutService}, \texttt{aiPlanService}, \texttt{profileService},
    etc.).
    \item \textbf{Library / Context} — Core utilities (Supabase client, Groq
    API wrapper, calculation functions, logger) and global state providers
    (Auth, Today, Fitness, Milestone).
\end{itemize}

This layered approach ensures that each layer only depends on the layer below
it, making the codebase maintainable, testable, and easy to extend.

\subsection{Database Schema}

The Supabase PostgreSQL database contains the key tables described in
Table~\ref{tab:schema}.

\begin{table}[h!]
\centering
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{|p{4.5cm}|p{9cm}|}
\hline
\textbf{Table} & \textbf{Purpose} \\
\hline
\texttt{profiles} &
    User profile: goal, body metrics, experience, equipment, diet, login
    streak \\
\hline
\texttt{calorie\_targets} &
    Computed daily calorie and macro targets (TDEE, protein, carbs, fat) \\
\hline
\texttt{training\_plans} &
    AI-generated weekly plan stored as JSON, one row per user (upsert) \\
\hline
\texttt{workout\_sessions} &
    Each workout session: start/end time, calories burned, notes \\
\hline
\texttt{workout\_exercises} &
    Individual exercises within a session: sets, reps, weight, posture score \\
\hline
\texttt{muscle\_fatigue} &
    Per-muscle fatigue percentage (0–100\%), updated after each workout \\
\hline
\texttt{food\_logs} &
    Each food logged: user reference, food reference, quantity in grams,
    timestamp \\
\hline
\texttt{foods} &
    Food database: calories, protein, carbs, fat per 100g \\
\hline
\texttt{daily\_activity} &
    Steps, sleep hours, and water intake per day per user \\
\hline
\texttt{chat\_messages} &
    Yara conversation history per user \\
\hline
\texttt{ai\_insights} &
    Cached AI insight cards with a 6-hour time-to-live \\
\hline
\texttt{direct\_messages} &
    Direct message conversations between users \\
\hline
\end{tabular}
\caption{Supabase Database Schema}
\label{tab:schema}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Development Support}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Development Tools}

The following tools were used throughout the development process:

\begin{itemize}
    \item \textbf{Visual Studio Code} — Primary code editor with React Native
    and TypeScript extensions.
    \item \textbf{Expo Go} — Mobile application for testing on physical devices
    without building native binaries.
    \item \textbf{Expo CLI} — Running the development server and managing
    builds.
    \item \textbf{Supabase Studio} — Web dashboard for managing the database,
    writing SQL queries, and deploying serverless Edge Functions.
    \item \textbf{Git \& GitHub} — Version control using a feature branch
    workflow; each feature was developed on a separate branch and merged to
    \texttt{main} after review.
    \item \textbf{Postman} — Testing Groq and Gemini API endpoints during
    integration.
    \item \textbf{Android Emulator / iOS Simulator} — Testing on virtual
    devices without requiring a physical phone.
\end{itemize}

\subsection{Version Control Strategy}

Development used a \textbf{feature branch workflow}:
\begin{itemize}
    \item \texttt{main} branch: stable, deployable code at all times.
    \item Feature branches named by feature
    (e.g., \texttt{mayssa-fix1}, \texttt{scheduleweek-fix-generate}).
    \item Branches are merged into \texttt{main} after review and testing.
\end{itemize}

This approach prevented conflicts when multiple team members worked on
different features simultaneously, and ensured the main branch always remained
functional.

\subsection{AI API Integration}

Two external AI providers are integrated into the application:

\subsubsection{Groq API}
\begin{itemize}
    \item Endpoint: \texttt{https://api.groq.com/openai/v1/chat/completions}
    \item Model \texttt{llama-3.1-8b-instant} — used for Yara's conversational
    chat, prioritizing low latency.
    \item Model \texttt{llama-3.3-70b-versatile} — used for AI workout plan
    generation, prioritizing output quality.
    \item The Yara Insights feature routes through a Supabase Edge Function
    so the API key is never exposed on the client side.
\end{itemize}

\subsubsection{Google Gemini API}
\begin{itemize}
    \item Endpoint: \texttt{generativelanguage.googleapis.com/v1beta/models/gemini-2.0-flash}
    \item Used in the onboarding plan generation flow.
    \item Chosen for its generous free tier (1 million tokens per minute),
    avoiding rate-limit issues during testing.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\section{Implementation}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{System Interfaces and Screens}
\label{sec:screens}

This section describes each screen of the application, its purpose, and its
key user interface elements.

\subsubsection{Authentication Screens}

The \textbf{Sign In} screen is the entry point for returning users. It
contains an email field, a password field with a show/hide toggle, and a
sign-in button that authenticates the user via Supabase Auth. The
\textbf{Sign Up} screen allows new users to register by providing their full
name, email, and password. On success, a corresponding row is created in the
\texttt{profiles} table and the user is redirected to the onboarding flow.

\subsubsection{Onboarding Flow}

New users complete a multi-step onboarding process that collects all
information needed to personalize the AI experience. The ten steps are:

\begin{enumerate}
    \item \textbf{Goal selection} — Lose Fat, Gain Muscle, Maintain, Gain
    Weight, Build Habits, or Athletic Performance.
    \item \textbf{Body measurements} — Height (cm), current weight (kg),
    target weight, and date of birth.
    \item \textbf{Experience level} — Beginner, Intermediate, or Advanced,
    each with a description of the associated training approach.
    \item \textbf{Equipment} — Full Gym, Dumbbells, Home (Bodyweight),
    Resistance Bands, or Kettlebells.
    \item \textbf{Schedule} — Number of workout days per week (1–7), session
    duration (20–90 min), and preferred time of day.
    \item \textbf{Focus areas} — Multi-select: Chest, Back, Legs, Arms,
    Shoulders, Core, Cardio.
    \item \textbf{Injuries and limitations} — Multi-select: None, Knee,
    Back, Shoulder, Wrist, Ankle.
    \item \textbf{Diet preference} — Anything, Vegetarian, Vegan, Keto,
    Gluten-Free, or Lactose-Free.
    \item \textbf{Lifestyle} — Sleep quality (1–5) and stress level (1–5).
    \item \textbf{Completion screen} — Animated confirmation before entering
    the main application.
\end{enumerate}

On completion, all data is saved to the \texttt{profiles} table, calorie
and macro targets are computed and saved to \texttt{calorie\_targets}, and
an interactive app tour launches to guide the user through the main tabs.

\subsubsection{Home Dashboard}

The home screen is the central view the user sees every day. It contains:
\begin{itemize}
    \item \textbf{Calorie Ring} — A circular animated progress indicator
    showing calories consumed versus the daily target.
    \item \textbf{Macro bars} — Horizontal progress bars for protein,
    carbohydrates, and fat.
    \item \textbf{Water tracker} — Tap-to-add water glasses with a
    hydration goal bar.
    \item \textbf{Today's Schedule Widget} — A preview of the day's
    exercises and meals from the AI plan, linking to the full Schedule screen.
    \item \textbf{Step Counter} — Real-time step count with daily goal
    progress from the device pedometer.
    \item \textbf{Quick action buttons} — Log Food, Log Sleep, Start
    Workout, Open Yara Chat.
\end{itemize}

The Yara floating action button is visible on this screen and provides
instant access to the AI coach from anywhere in the app.

\subsubsection{Training Screen}

The training hub allows the user to:
\begin{itemize}
    \item Generate a personalized AI workout plan by tapping the
    \textbf{Generate Plan} button.
    \item Regenerate with a different training split style (Push/Pull/Legs,
    Upper/Lower, Full Body, Bro Split, Athletic, or Hypertrophy).
    \item Browse the exercise library with category filters.
    \item Access the PostureAI camera for real-time form analysis.
    \item View workout history.
\end{itemize}

\subsubsection{Schedule Screen}

This screen displays the AI-generated weekly plan interactively:
\begin{itemize}
    \item A \textbf{horizontal day-selector} with colored pills for each
    day, indicating workout type, completion status, and missed days.
    \item A \textbf{progress card} showing workout completion percentage,
    step count versus goal, and planned meal calories.
    \item An \textbf{exercise checklist} with animated rows; exercises
    carried over from a missed previous day are tagged ``carried over''.
    \item \textbf{Meal cards} showing each planned meal with per-food
    calorie breakdown.
    \item A \textbf{Mark Day Complete} button that activates only when
    every exercise is checked off.
\end{itemize}

\subsubsection{Workout Screens}

The \textbf{WorkoutActive} screen provides the live workout interface with
rep counting and a rest timer. The \textbf{WorkoutSummary} screen displays
post-workout statistics including total reps, calories burned, duration, and
average posture score. The \textbf{PostureAI} screen uses the device camera
with a real-time pose estimation model to provide visual feedback on exercise
form. The \textbf{FlappyBird Game} is a mini-game that activates during rest
periods to keep users engaged.

\subsubsection{Workout History Screen}

A timeline of all past sessions, grouped by month. Each session card shows
exercise names, date, total reps, duration, calories burned, and a color-coded
FORM score (green $\geq$80\%, orange $\geq$55\%, red below). Tapping a card
opens a detail sheet with a full exercise breakdown showing sets, reps, weight,
duration, and posture score per exercise.

\subsubsection{Nutrition Screens}

The \textbf{Nutrition} overview shows the day's calorie ring and food log
entries. The \textbf{Meal Logger} provides three methods to log food:
text search of the local food database, barcode scanning via the device
camera, and AI photo analysis. The \textbf{Custom Meal Builder} allows users
to create named meal presets from multiple ingredients and log them in a
single tap.

\subsubsection{Insights Screen}

The analytics dashboard contains:
\begin{itemize}
    \item A \textbf{period selector} (Week / Month / 3 Months).
    \item \textbf{Four summary cards}: Streak (days), Workouts (count),
    Average Steps, and Average Sleep.
    \item A \textbf{Health Score Chart} showing daily scores (0–100) over
    the selected period.
    \item A \textbf{42-day activity heatmap} showing intensity per day.
    \item \textbf{Nutrition and Workout summaries} with averaged metrics.
    \item A \textbf{Muscle Fatigue chart} showing recovery state per muscle
    group.
    \item \textbf{Four AI Insight Cards} generated by Yara from the user's
    real behavioral data (cached for 6 hours).
\end{itemize}

\subsubsection{Community Screens}

The \textbf{Community Center} is a public social feed for posting workout
achievements. The \textbf{Messages Inbox} lists all direct message
conversations. The \textbf{DM Thread} provides real-time 1-to-1 messaging
with push notification support for new messages received in the background.

\subsubsection{Profile and Settings}

The \textbf{Profile} screen displays the user's avatar, name, bio, location,
and activity summary. The \textbf{Edit Profile} screen allows updating the
display name (required), bio (160 characters maximum), date of birth
(native date picker), city, country (searchable list of 70+ countries), and
password. Settings sub-screens include Help Center, Report a Problem, Terms
and Policies, and Trust Center.

% ─────────────────────────────────────────────────────────────────────────────

\subsection{Feature Implementation}
\label{sec:features}

\subsubsection{Authentication and Session Management}

Authentication is handled by Supabase Auth. The \texttt{AuthContext}
provider wraps the entire application and manages the global user state.
On every app start, it calls \texttt{supabase.auth.getSession()} and
subscribes to \texttt{supabase.auth.onAuthStateChange()} to react
automatically to login and logout events.

When a user logs in, the following sequence executes:
\begin{enumerate}
    \item The user's training plan is fetched from Supabase and loaded into
    the local \texttt{scheduleStore}.
    \item If no Supabase plan exists, the local AsyncStorage is checked as
    a fallback.
    \item The user's profile is fetched to determine whether onboarding is
    complete (\texttt{profile.goal \&\& profile.onboarded}).
    \item A server-side RPC call (\texttt{record\_user\_visit}) updates the
    login streak in the \texttt{profiles} table.
    \item The local avatar is loaded from device storage.
\end{enumerate}

Navigation is determined by the following logic:
\begin{itemize}
    \item No active session $\rightarrow$ Sign In / Sign Up screens.
    \item Session exists, onboarding incomplete $\rightarrow$ Onboarding
    flow.
    \item Session exists, onboarding complete $\rightarrow$ Main application
    (five-tab navigation).
\end{itemize}

On logout, \texttt{scheduleStore.clear()} wipes the local plan from the
device to prevent data leakage between accounts.

\subsubsection{Onboarding and User Profiling}

On completion of the onboarding form, the \texttt{calculations.ts} library
computes the user's energy requirements using the Mifflin-St Jeor equations:

\begin{align}
    \text{BMR (male)} &= 10w + 6.25h - 5a + 5 \\
    \text{BMR (female)} &= 10w + 6.25h - 5a - 161
\end{align}

where $w$ is weight in kg, $h$ is height in cm, and $a$ is age in years.

The TDEE (Total Daily Energy Expenditure) is then computed as:
\begin{equation}
    \text{TDEE} = \text{BMR} \times \text{activity multiplier}
\end{equation}

The activity multiplier ranges from 1.2 (sedentary) to 1.725 (very active).
The daily calorie target is adjusted based on the user's goal:
\begin{itemize}
    \item Lose fat: TDEE $-$ 500 kcal/day
    \item Gain muscle: TDEE $+$ 250 kcal/day
    \item Gain weight: TDEE $+$ 500 kcal/day
    \item Maintain: TDEE
\end{itemize}

Macro targets are distributed as follows: protein at 2g per kg of body
weight (muscle gain) or 1.8g (fat loss), fat at 25\% of total calories, and
carbohydrates filling the remaining calorie budget.

\subsubsection{AI Workout Plan Generation}

The plan generation feature follows a five-step agentic workflow:

\begin{enumerate}
    \item \textbf{Profile retrieval} — Two Supabase queries run in parallel
    using \texttt{Promise.all()}: one fetching the user's profile and one
    fetching their calorie targets.

    \item \textbf{Prompt engineering} — The \texttt{buildPrompt()} function
    maps raw profile fields to fitness-specific instructions. For example,
    the experience level maps as follows:
    \begin{itemize}
        \item Beginner $\rightarrow$ ``simple compound movements, reps 12–15''
        \item Intermediate $\rightarrow$ ``compound + isolation, reps 8–12''
        \item Advanced $\rightarrow$ ``heavy compounds, reps 4–10, periodisation''
    \end{itemize}
    A random seed number is prepended to every prompt to ensure each
    generation produces a unique plan.

    \item \textbf{Groq API call} — The prompt is sent to
    \texttt{llama-3.3-70b-versatile} with \texttt{max\_tokens: 4096},
    \texttt{temperature: 0.8}, and \texttt{response\_format: json\_object}
    to enforce structured JSON output.

    \item \textbf{Defensive parsing} — The \texttt{parsePlanResponse()}
    function repairs common AI formatting errors: stripping markdown fences,
    removing trailing commas, fixing unquoted keys, removing control
    characters, and auto-closing any unclosed braces or brackets. A second
    parse attempt is made after repair before throwing an error.

    \item \textbf{Persistence} — The parsed plan is saved to
    \texttt{training\_plans} via an upsert operation with
    \texttt{onConflict: 'user\_id'}, ensuring each user has exactly one
    active plan. The plan is simultaneously written to the local
    \texttt{scheduleStore} so the Schedule screen updates without a reload.
\end{enumerate}

Regeneration cycles through six predefined training split styles using a
modulo index:
\[
\text{nextIndex} = (\text{currentIndex} + 1) \mod 6
\]
This guarantees structural variety (not just different exercise names) across
regenerations.

\subsubsection{Yara AI Chat Assistant}

Yara is the application's central conversational AI coach. Its system prompt
defines a personality (warm, direct, coach-like, no AI stiffness) and injects
the user's full profile — goal, weight, experience, equipment, injuries,
daily calories, and macros — at every call. Yara is instructed never to ask
for information already present in the profile.

The \texttt{useYaraChat} hook implements two output modes:

\begin{itemize}
    \item \textbf{Normal chat mode:} The user's message is sent to Groq
    \texttt{llama-3.1-8b-instant} with a 512-token limit for fast responses.
    The reply is displayed as a chat bubble and saved to
    \texttt{chat\_messages}.

    \item \textbf{Schedule generation mode:} Triggered by keyword detection
    (e.g., ``give me a weekly plan'', ``workout plan'', ``plan my week'').
    A strict JSON schema is appended to the user's message. The model
    responds with a full 7-day plan in JSON format, which is parsed and
    saved both to \texttt{scheduleStore} (for immediate UI update) and to
    the \texttt{training\_plans} table (for cross-device persistence).
\end{itemize}

Conversation history is persisted to Supabase and reloaded each time the
chat opens, allowing the model to maintain context across sessions.

\subsubsection{Schedule Store and Completion Tracking}

The \texttt{scheduleStore} is a custom lightweight pub/sub store that manages
the weekly plan state and exercise completion. Its state is structured as
follows:

\begin{itemize}
    \item \textbf{Schedule state:} The full AI plan JSON (days, exercises,
    meals, step and sleep targets).
    \item \textbf{Completion state:} A date-keyed map recording which
    exercises have been checked off and whether the day is marked done,
    e.g., \texttt{\{ "2026-05-13": \{ done: false, checked: [0, 2] \} \}}.
\end{itemize}

All writes are immediately persisted to AsyncStorage, making the schedule
available offline. Components subscribe to the store and re-render
automatically on state changes, without requiring a page reload.

Two adaptive behaviors are implemented in the Schedule screen:
\begin{itemize}
    \item \textbf{Exercise carryover:} If yesterday's workout was not
    completed, its uncompleted exercises are appended to today's list, tagged
    as ``carried over''.
    \item \textbf{Adaptive step targets:} If the user fell short of their
    step target the previous day, today's target is increased by the
    shortfall.
\end{itemize}

\subsubsection{Workout Tracking and Muscle Fatigue}

When a workout session ends, \texttt{saveWorkoutSession()} performs two
parallel database operations:

\begin{enumerate}
    \item \textbf{Session insertion} — A row is inserted into
    \texttt{workout\_sessions} with start/end timestamps, calories burned,
    and a notes summary (e.g., ``Squat · 12 reps · 78\% form'').

    \item \textbf{Muscle fatigue update} — Each exercise maps to a set of
    muscle groups (e.g., squat $\rightarrow$ Quads, Glutes, Hamstrings).
    For each targeted muscle, the stored fatigue percentage is increased by
    20 percentage points, capped at 100\%:
    \[
        \text{fatigue}_{\text{new}} = \min(100,\; \text{fatigue}_{\text{current}} + 20)
    \]
\end{enumerate}

Muscle fatigue data is displayed in the Insights screen and is also included
in the context sent to the Yara Insights AI, enabling recovery-aware
recommendations.

\subsubsection{Insights and AI Analysis}

The Insights screen aggregates data from five sources simultaneously using
\texttt{Promise.all()}: a Supabase RPC returning aggregated stats and a
42-day activity heatmap, direct food log queries for macro breakdowns, direct
workout session queries for training summaries, muscle fatigue records, and
cached AI insight history.

The \textbf{Health Score} formula assigns a daily score from 0 to 100:
\begin{align*}
    \text{Workout completed} &\rightarrow +35 \\
    \text{Steps} > 8{,}000  &\rightarrow +25 \quad (\text{steps} > 4{,}000 \rightarrow +12) \\
    \text{Food logged}       &\rightarrow +20 \\
    \text{Sleep} \geq 7\text{h} &\rightarrow +15 \quad (\text{sleep} \geq 6\text{h} \rightarrow +8) \\
    \text{Water} \geq 2\text{L} &\rightarrow +5 \\
    \hline
    \text{Maximum}           &= 100
\end{align*}

The \textbf{AI Insight Cards} are generated through a server-side Supabase
Edge Function to protect the Groq API key. A 6-hour cache is implemented
in the \texttt{ai\_insights} table: if four fresh rows exist, they are
returned immediately without calling the AI; otherwise, the Edge Function
is invoked, new cards are generated and stored, then returned to the screen.
A deduplication map prevents multiple concurrent requests from each
independently triggering a cache miss and calling the API redundantly.

% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
% ─────────────────────────────────────────────────────────────────────────────

This chapter has presented the complete development and prototype of the
BodyQ mobile application. The implementation demonstrates that a
sophisticated, AI-powered fitness application can be built using modern
cross-platform tools while maintaining high code quality, a clean
architecture, and a smooth user experience.

The key achievements of this development phase are:

\begin{enumerate}
    \item \textbf{A complete agentic AI system} — Yara functions as a
    genuine AI coach: she understands the user's profile, generates
    structured weekly plans from natural language requests, and produces
    analytical insights derived from real behavioral data, all without
    requiring the user to understand the underlying technology.

    \item \textbf{Personalization at every level} — From the calorie ring
    on the home screen to the specific exercises in the weekly plan, every
    piece of content is derived from the individual user's profile. No two
    users see the same plan.

    \item \textbf{Robust offline capability} — The weekly schedule and
    exercise completion state persist locally on the device, meaning the
    core tracking features function without an internet connection.

    \item \textbf{Scalable, maintainable architecture} — The layered
    architecture (Screens $\rightarrow$ Hooks $\rightarrow$ Services
    $\rightarrow$ Lib) isolates each feature, making the codebase testable
    and replaceable without affecting the rest of the system.

    \item \textbf{Production-ready AI integration} — The application
    demonstrates practical AI engineering: structured prompt design,
    defensive JSON parsing, server-side API key protection via Edge
    Functions, intelligent result caching, and concurrent-request
    deduplication.
\end{enumerate}

The prototype is fully functional and covers all core user journeys defined
in the requirements. Future development will focus on expanding the exercise
library, refining the AI models with fitness-specific data, and introducing
social features such as group challenges and leaderboards.
