% CHAPTER 5 — GENERAL CONCLUSION

\chapter{General Conclusion \& Future Work}

\section{Project Review}

BodyQ was conceived to address a critical gap in the digital fitness
landscape: existing applications collect health data without genuinely
acting on it, offer static workout templates devoid of real
personalisation, and provide no real-time feedback on exercise technique.
Over nine weeks and five Agile Scrum sprints, a team of five students
designed, implemented, and validated a cross-platform AI-powered fitness
and health assistant that integrates personalisation, intelligence, and
usability into a single coherent product.

\section{Achievement of Objectives}

The four core objectives stated in Chapter~1 were addressed as follows.

\begin{enumerate}
	\item \textbf{Personalisation.} A ten-step onboarding flow collects
	goals, body metrics, experience level, available equipment, preferred
	schedule, dietary restrictions, and lifestyle indicators. These inputs
	feed a Supabase Edge Function that invokes the Groq Llama~3.1-8B
	model to generate a personalised multi-week training plan with
	calibrated daily calorie and macronutrient targets derived from BMR
	and TDEE calculations.
	
	\item \textbf{Real-time form feedback.} The posture detection module
	integrates MediaPipe Pose via a WebView bridge to compute joint angles
	in real time during workout sessions. The system tracks repetition
	counts and delivers live correction cues displayed as overlays on the
	camera feed.
	
	\item \textbf{Cross-domain data integration.} The Yara AI coaching
	engine (Groq Llama~3.3-70B) aggregates 30-day activity history,
	nutrition logs, body metrics, and previous coaching interactions into
	a single RAG-augmented prompt context before generating each response.
	Semantic retrieval of embedded user data via OpenAI
	\texttt{text-embedding-3-small} further enriches context with
	historically relevant health events.
	
	\item \textbf{Behavioral insights.} The Insights screen produces
	AI-generated analytics cards via the \texttt{yara-insights} Edge
	Function, categorised by tag (Performance, Nutrition, Recovery,
	Correlation, Optimisation, or Prediction) and parameterised by the
	user's real activity statistics for the selected reporting period.
\end{enumerate}

Beyond the stated objectives, the project delivered a voice interface
enabling hands-free coaching queries and data logging via OpenAI/Groq
Whisper speech-to-text; a gamification layer comprising XP points, login
streaks, and unlockable achievement badges; push notifications for workout
and hydration reminders; and a Next.js~14 administration dashboard for
monitoring user registrations and AI interaction logs.

\section{Technical Challenges}

Several non-trivial engineering challenges were encountered and resolved
during development.

\textbf{MediaPipe integration.} No production-grade native React Native
binding for MediaPipe Pose was available at the project start. The solution
was to embed the pose estimation pipeline inside a React Native WebView
with a bi-directional JavaScript messaging interface, allowing the mobile
application to receive per-frame joint angle and repetition data from the
web-based MediaPipe runtime.

\textbf{LLM output reliability.} Long AI prompts aggregating multiple data
sections occasionally produced malformed JSON from the Groq API. A
defensive repair pipeline---removing trailing commas, stripping control
characters, and balancing unclosed brackets---was implemented to guarantee
robust parsing without degrading the user experience.

\textbf{Row Level Security and Edge Functions.} Supabase RLS policies that
correctly restrict mobile clients to their own data must simultaneously
permit Edge Functions---which authenticate with a service-role key---to
read and write across all user records. Calibrating this boundary required
careful policy design to avoid either over-permitting mobile access or
blocking legitimate server-side operations.

\textbf{Offline-first behaviour.} Network requests from the mobile
application are not guaranteed to succeed. AsyncStorage caching of profile
data and recent dashboard statistics ensures core functionality remains
accessible when connectivity is unavailable, and TanStack Query's
stale-while-revalidate strategy keeps the UI responsive during background
refreshes.

\section{Future Work}

The BodyQ prototype establishes a strong technical foundation. The
following extensions are identified as priorities for future development.

\begin{enumerate}
	\item \textbf{Wearable integration.} Sprint~5 included a preliminary
	exploration of wearable device connectivity. A production-quality
	integration with Apple HealthKit, Google Fit, or Bluetooth Low Energy
	(BLE) fitness trackers would enable passive step counting, continuous
	heart rate monitoring, and automatic workout detection without manual
	logging.
	
	\item \textbf{Native posture detection.} Migrating the MediaPipe
	pipeline from the WebView sandbox to a native Expo module (via the
	Expo Modules API) would eliminate the WebView communication overhead,
	improve frame processing latency, and allow full-screen camera overlays
	during workout sessions.
	
	\item \textbf{Social and community features.} The community screen was
	scaffolded but not completed within the project timeline. Future
	iterations could implement workout sharing, friend-challenge systems,
	and group streak competitions to reinforce behavioural adherence
	through social accountability.
	
	\item \textbf{Cycle-aware coaching.} The data model and AI coaching
	architecture already include support for menstrual cycle information.
	A dedicated cycle-tracking interface and cycle-phase-sensitive coaching
	prompts---adapting workout intensity and nutritional guidance to the
	follicular, ovulatory, luteal, and menstrual phases---are planned as
	a significant differentiator for female users.
	
	\item \textbf{Clinical and user-study validation.} A structured study
	with fitness professionals and healthcare practitioners is needed to
	validate the accuracy of the posture detection system, the
	appropriateness of AI coaching responses, and the usability of the
	onboarding flow across diverse user populations.
	
	\item \textbf{On-device AI inference.} Deploying a quantised language
	model on-device would allow the coaching engine to function without
	internet connectivity, addressing the current dependency on Groq cloud
	inference for all AI features.
\end{enumerate}

\section{Conclusion}

BodyQ demonstrates that a compact academic team can deliver a technically
ambitious, AI-integrated health application within a constrained nine-week
timeline by combining modern cloud-native infrastructure (Supabase),
capable open-source large language models (Llama~3.3 via Groq), and
cross-platform mobile tooling (React Native with Expo~SDK~54). The
resulting prototype validates the core hypothesis of the project: that
genuine personalisation, real-time biomechanical feedback, and behavioral
correlation analysis can coexist in a single consumer-grade fitness
application without requiring proprietary hardware. The architecture and
codebase established through this project provide a robust platform for
the clinical validation, feature expansion, and community building
identified as the next steps toward a production-ready product.
