% CHAPTER 3 — PRODUCT DESIGN

\chapter{Product Design}

\section{Introduction}
This chapter presents the complete design of the BodyQ system at two
levels of abstraction. The high-level design section describes the
overall system architecture and the decomposition of the platform into
its major components. The detailed design section presents the key
interaction flows as sequence diagrams, the process models as activity
diagrams, and the data model as a class/entity description. The chapter
concludes with an overview of the user interface design.

\section{High-Level Design}

\subsection{Architecture Overview}
BodyQ is structured as a four-tier distributed system consisting of a
mobile client, a backend-as-a-service layer, a serverless AI engine
layer, and a web administration interface.

\textbf{Tier~1 --- Mobile Client (React Native / Expo SDK~54).}
The mobile application is built with React Native and Expo SDK~54 and
targets iOS~16+ and Android~11+. It is organized using the Expo Router
file-based navigation system. Application state is managed with Zustand
for global stores (authentication, user profile, active workout session)
and TanStack Query for server state and cache management. Forms are
handled with React Hook Form and validated using Zod schema definitions.

\textbf{Tier~2 --- Backend-as-a-Service (Supabase).}
Supabase provides the entire backend infrastructure: a managed
PostgreSQL~15 database, an authentication service (email/password and
Google OAuth~2.0), object storage for meal photos and profile images,
and Edge Functions running on the Deno runtime. Row Level Security
policies are applied to every table in the schema to enforce user data
isolation at the database level.

\textbf{Tier~3 --- AI Engine Layer.}
The AI engine is implemented as five Supabase Edge Functions:

\texttt{ai-assistant} (RAG conversational coaching)

\texttt{daily-tip} (daily AI insights)

\texttt{analyze-food} (meal photo recognition)

\texttt{generate-workout-plan} (personalized training programs)

\texttt{recommendation-engine} (rule-based recommendations).

These functions invoke the Claude API (Anthropic), the Groq inference platform(\texttt{llama-3.1-8b-instant}) or (\texttt{llama-3.3-70b-instant}), or the Google Gemini Vision API
depending on the task requirements and latency constraints.

\textbf{Tier~4 --- Administration Dashboard (React 18.0, Next.js~14).}
A web-based administration dashboard built with React 18.0, Next.js~14 (App Router) and Tailwind CSS provides operational management. 
All Supabase calls from the dashboard use the service role key exclusively via server-side Route Handlers.

\subsection{Component Diagrams}
% TODO: UML Component diagram

\section{Detailed Design}

\subsection{Sequence Diagrams}
\textbf{SD-01: User Registration and AI-Personalized Onboarding.}

\begin{enumerate}
	\item User submits the registration form; mobile client calls
	\texttt{supabase.auth.signUp()}.
	\item Supabase Auth creates the user record and returns a session JWT.
	\item Mobile client navigates to onboarding; user completes all
	5~steps; collected data is buffered in a Zustand
	\texttt{onboardingStore}.
	\item On step~5 completion, the mobile client calls the
	\texttt{generate-workout-plan} Edge Function with the user profile
	JSON.
	\item The Edge Function invokes the Claude API and receives a
	personalized 4-week workout plan as structured JSON.
	\item The Edge Function computes BMR and TDEE; inserts records into
	\texttt{profiles} and \texttt{calorie\_targets}.
	\item The mobile client stores the plan and navigates to the Home tab.
\end{enumerate}

\textbf{SD-02: RAG Coaching Pipeline (Yara AI Assistant).}

\begin{enumerate}[noitemsep]
	\item User types a question in the AI assistant chat screen.
	\item Mobile client calls
	\texttt{supabase.functions.invoke('ai-assistant')} with
	\texttt{\{userId, query\}}.
	\item \textbf{Retrieve:} Edge Function executes four SQL queries---7-day
	\texttt{daily\_activity} window, 7-day \texttt{food\_logs}
	aggregated by day, active \texttt{calorie\_targets}, and last~5
	\texttt{workout\_sessions}.
	\item \textbf{Augment:} Retrieved records are formatted into a
	structured plain-text context block; out-of-range values (e.g.,
	sleep $<6$\,h, hydration $<1500$\,mL) are flagged with warning
	markers; four behavioral rules are appended.
	\item \textbf{Generate:} Augmented prompt is sent to the Groq API
	(\texttt{Llama3.3 70b}).
	\item Edge Function inserts the response into \texttt{ai\_insights}
	(\texttt{type='chat'}) and returns it to the mobile client.
\end{enumerate}

\textbf{SD-03: Posture Detection During Workout.}

\begin{enumerate}[noitemsep]
	\item User taps ``PostureAI'' during an active workout session.
	\item Mobile client renders \texttt{PostureScreen} with an Expo
	WebView component.
	\item WebView HTML loads MediaPipe Pose JS from CDN, initializes the
	model, and starts the device camera via \texttt{getUserMedia}.
	\item At each video frame, MediaPipe detects 33~landmarks; the array
	is serialized to JSON and posted to React Native via
	\texttt{window.ReactNativeWebView.postMessage()}.
	\item \texttt{PostureEngine.analyzeSquat(landmarks)} is called; joint
	angles are computed using the vector dot-product formula and
	evaluated against reference ranges; rep direction changes are
	tracked.
	\item Score, feedback strings, and rep count are rendered on the
	overlay UI in real time.
\end{enumerate}

\subsection{Activity Diagrams}
% TODO: Workout session flow, onboarding flow

\subsection{Class Diagram}
% TODO: Based on DB schema

\section{UI Design}
% TODO: Wireframes / mockups

\section{Conclusion}
